\documentclass[12pt,letterpaper]{article}
\usepackage{inverba}
\newcommand{\userName}{Cullyn Newman} 
\newcommand{\class}{BI 216} 
\newcommand{\institution}{Portland State University} 
\newcommand{\thetitle}{\hypertarget{home}{Lab 3 Addendum: Water Quality}}
\rfoot{\hyperlink{home}{\thepage}}

\begin{document}

\section*{Part A: Smithsonian Ocean}
\begin{enumerate}[font=\bfseries, wide]
    {\color{under}\item What does the figure just below the  “A More Acidic Ocean” figure illustrate? Please summarize the data in the figure (remember to account for all of the variables!) \textbf{(2pts)}}
    {\color{under}\item What does “ocean acidification” mean? \textbf{(1 pt)}}
    {\color{under}\item How does the level of acidity in the ocean impact:}
    \begin{enumerate}
        {\color{under}\item  A bivalve’s shell development? \textbf{(0.5pts)}}
        {\color{under}\item  A coral’s structure? \textbf{(0.5pts)}}
    \end{enumerate}
    {\color{under}\item Follow the links, and work to find the actual article on the research on black-finned clownish which was described in the In the Lab section}
    \begin{enumerate}
        {\color{under}\item What is the name of the journal in which it was published, and what was the (print) publication date and year \textbf{(0.25pts)}}
        {\color{under}\item The study was novel because it showed that fish have the ability to respond to predator sounds when played underwater:  TRUE or FALSE \textbf{(0.25pts)}}
    \end{enumerate}
\end{enumerate}

\section*{Part B: Measuring Salinity in Estuaries}
\subsection*{Level 3: Measuring Salinity in Estuaries}
\begin{enumerate}[font=\bfseries, wide, resume]
    {\color{under}\item Question \#5 from the activity: Which statement represents a valid conclusion based on the graph?  Enter the correct letter and the statement \textbf{(0.5pts)}}

    C. A rainstorm on Oct 25 may have caused the decrease in salinity on Oct 27

    {\color{under}\item What may have caused Delta Smelt to be found outside of their normal range? \textbf{(0.5pts)} }

    There was a significant amount of rainfall during the times the salinity was higher than 2, and salinity also spiked when ever rainfall increased. 
\end{enumerate}

\subsection*{Level 4 - Research Question: Predicting the Return of the Atlantic Sturgeon}
\begin{enumerate}[font=\bfseries, wide, resume]
    {\color{under}\item To get started, use the online Fact Sheet to select an estuary where Atlantic Sturgeon are found. Record the estuary name and location here:  \textbf{(0.5pts)}  }

    The location we chose was Chesapeake Bay, MD. 

    {\color{under}\item Write your research question in the space below. \textbf{(1pt)} }

    How does dissolved oxygen and temperature change over 

    {\color{under}\item Complete the table \textbf{(1pt)}}
    {\color{under}\item Can you identify a time period when the water temperature is within the range for the sturgeon to return? \textbf{(0.5pts)}}
    {\color{under}\item What is the range of the other water quality parameters during that time period? \textbf{(0.5pts)}}
    {\color{under}\item Can you identify a time period when all the conditions look right for the sturgeon to return to spawn? \textbf{(0.5pts)}}
    {\color{under}\item Do the same conditions occur around the same time, year after year? \textbf{(0.5pts)}}
\end{enumerate}

\subsection*{Level 5: Work as a team to develop your own investigation}
\begin{enumerate}[font=\bfseries, wide, resume]
    {\color{under}\item Read through Level 5 on your own, and then work with your team to develop your research question. State your research question here: \textbf{(1 pt)}}
    {\color{under}\item State your hypothesis: \textbf{(1 pt)}}
    {\color{under}\item \textit{Make a Plan:} Make a lis below of the specific data you will need to answer the question \textbf{(1 pt)} }
    {\color{under}\item Other than the data listed above, what other information (if any) will you need to answer your question? \textbf{(1 pt)}}
    {\color{under}\item Insert figure here \textbf{(1 pt)}}
    {\color{under}\item \textit{Interpret the data:} What does your data show? Be specific and descriptive. Does the data support your hypothesis? \textbf{(1 pt)}}
    {\color{under}\item \textit{Draw a Conclusion:} What is the answer to your question? Use evidence and data to support your conclusion. \textbf{(1 pt)}}
    {\color{under}\item Give a specific example of why it would be biologically relevant to measure \textbf{temperature} in an aquatic environment. \textbf{(1 pt)}}
    {\color{under}\item Give a specific example of why it would be biologically relevant to measure \textbf{dissolved oxygen} in an aquatic environment. \textbf{(1 pt)}}
    {\color{under}\item Give a specific example of why it would be biologically relevant to measure \textbf{carbon dioxide} in an aquatic environment. \textbf{(1 pt)}}
\end{enumerate}
\end{document}