\documentclass[12pt,a4paper]{article}
\usepackage{inverba}
\newcommand{\userName}{Cullyn Newman} 
\newcommand{\class}{BI 216} 
\newcommand{\institution}{Portland State University} 
\newcommand{\thetitle}{\hypertarget{home}{Lab 4: Diffusion, Osmosis, Tonicity}}
\rfoot{\hyperlink{home}{\thepage}}

\begin{document}
\section*{Part A: Diffusion}
\begin{enumerate}[font=\bfseries, wide]
    {\color{under}\item Describe two factors that could speed up the rate of diffusion. \textbf{(0.5 pt)}}

    {\color{under}\item At what point does net diffusion end? \textbf{(0.5 pt)}}

    >> Answer

    {\color{under}\item  The video discusses five factors that can affect the rate of diffusion. List one and explain it using your own words. \textbf{(1 pt)}}

    >> Answer
    
    {\color{under}\item Identify a biological situation in which efficient diffusion of a solute from one region to another would be a matter of life or death for an organism. Please describe that situation, and what factors would influence the diffusion. \textbf{(1 pt)}}

    >> Answer

\end{enumerate}

\section*{Part B: Osmosis and Tonicity; Simulation}
\begin{enumerate}[font=\bfseries, wide, resume]
    {\color{under}\item Explain one way that osmosis differs from diffusion, and one wya that it is similar. \textbf{(1 pt)}}

    >> Answer

    {\color{under}\item How long will the dialysis tubes remain in each beaker? Why does this matter? \textbf{(1 pt)}}

    >> Answer

    \begin{table}[h]
        \centering
        \caption{Caption}
        \begin{tabular}{r|ccccc}
            \toprule
            Trial & A & B & C & D & E\\
            \midrule
            Beaker \par \% Sugar &&&&\\
            Dialysis Tube\par \% sugar&&&&\\
            Initial Mass (g)&&&&\\
            Final Mass (g) &&&&\\
            $\Delta$ Mass (g)&&&&\\
            \bottomrule
            \end{tabular}
    \end{table}
    

    {\color{under}\item Name two variables that remained constant throughout the lab. \textbf{(1 pt)}}

    >> Answer

    {\color{under}\item Name the \textit{dependent} variable. \textbf{(1 pt)}}

    >> Answer

    {\color{under}\item Name the \textit{independent} variable.\textbf{(1 pt)}}

    >> Answer

    {\color{under} The percent concentration of both fluids is different. There is {\color{darkmodetext}\textbf{}\%} sugar in the beaker and {\color{darkmodetext}\textbf{}\%} sugar in the dialysis tube. Now, think of it from the water's point of  view. There is {\color{darkmodetext}\textbf{}\%} water in the beaker and {\color{darkmodetext}\textbf{}\%} in the dialysis tube}

    {\color{under}\item Answer the following:}
    \begin{enumerate}
        {\color{under}\item Did the beaker have a higher/lower concentration of water than the dialysis tube? \textbf{0.5 pt}}

        >> Answer 

        {\color{under}\item Did the water flow in or out of the tube and what type of diffusion is this? \textbf{0.5 pt}}

        >> Answer 
    \end{enumerate}

    {\color{under}\item Answer the following:}
    \begin{enumerate}
        {\color{under}\item Which beaker had a higher concentration of sugar solution fluid outside than inside? \textbf{0.5 pt}}

        >> Answer 

        {\color{under}\item How will the water flow in this situation? If the tube was a cell, what type of solution was the cell placed in? \textbf{0.5 pt}}

        >> Answer 
    \end{enumerate}

    {\color{under}\item Answer the following:}
    \begin{enumerate}
        {\color{under}\item Which beaker had a percent concentration that was equal on both inside/outside the cell? \textbf{0.5 pt}}

        >> Answer 

        {\color{under}\item How will the water flow in this situation? What type of solution was the cell (tube) placed in? \textbf{0.5 pt}}

        >> Answer 
    \end{enumerate}

    {\color{under}\item Answer the following:}
    \begin{enumerate}
        {\color{under}\item Which beaker has a higher concentration of sugar solution inside the cell than outside?  \textbf{0.5 pt}}

        >> Answer 

        {\color{under}\item How will the water flow in this situation? What type of solution was the cell (tube) placed in? \textbf{0.5 pt}}

        >> Answer 
    \end{enumerate}

    {\color{under}\item Graphically represent the results of today’s simulation.  \textbf{(2 pts)}}

    >> Answer

\end{enumerate}
    
\section*{Part C: Regulatory Mechanisms}
\begin{enumerate}[font=\bfseries, wide, resume]
    {\color{under}\item Describe one challenge inherent to being a marine organism.  \textbf{(1 pt)}}

    >> Answer

    {\color{under}\item Describe one challenge inherent to being a freshwater organism.  \textbf{(1 pt)}}

    >> Answer

    {\color{under}\item What mechanisms have evolved to compensate for these challenges for both types of organisms?  \textbf{(1 pt)}}

    >> Answer

    {\color{under}\item Describe mechanisms that have evolved to facilitate organisms that switch from freshwater to marine environments. \textbf{(2 pts)}}

    >> Answer

\end{enumerate}

\end{document}