\chapter{Week 8: Endocrine, Male Reproductive}

\section{Endocrine Glands}

\subsection{Pituitary Gland}
\begin{center}
  % \Image{0.94\columnwidth}{images/week-8-a.jpg}
\end{center}
\begin{itemize}
  \item 
\end{itemize}

\subsection{Adrenal Gland}
\begin{center}
  % \Image{0.94\columnwidth}{images/week-8-b.jpg}
\end{center}
\begin{itemize}
  \item 
\end{itemize}

\subsection{Questions}
\begin{itemize}\color{minor}
  \item What hormone is produced by the zona glomerulosa?
  \basec{\begin{itemize}
    \item 
  \end{itemize}}

  \item By the zona fasciculata?
  \basec{\begin{itemize}
    \item 
  \end{itemize}}
  
  \item By the zona reticularis?
  \basec{\begin{itemize}
    \item 
  \end{itemize}}
  
  \item By the medulla?
  \basec{\begin{itemize}
    \item 
  \end{itemize}}
  
  \item What hormones regulate the function of the cortex?
  \basec{\begin{itemize}
    \item 
  \end{itemize}}
  
  \item How is medullary function regulated?
  \basec{\begin{itemize}
    \item 
  \end{itemize}}

\end{itemize}

\subsection{Thyroid}
\begin{center}
  % \Image{0.94\columnwidth}{images/week-8-c.jpg}
\end{center}
\begin{itemize}
  \item 
\end{itemize}

\section{Male Reproductive System}

\subsection{Testis}
\begin{center}
  % \Image{0.94\columnwidth}{images/week-8-d.jpg}
\end{center}
\begin{itemize}
  \item 
\end{itemize}

\subsection{Epididymis}
\begin{center}
  % \Image{0.94\columnwidth}{images/week-8-e.jpg}
\end{center}
\begin{itemize}
  \item 
\end{itemize}

\subsection{Ductus Deferens}
\begin{center}
  % \Image{0.94\columnwidth}{images/week-8-k.jpg}
\end{center}
\begin{itemize}
  \item 
\end{itemize}

\subsection{Seminal Vesicle}
\begin{center}
  % \Image{0.94\columnwidth}{images/week-8-l.jpg}
\end{center}
\begin{itemize}
  \item 
\end{itemize}

\subsection{Prostate Gland}
\begin{center}
  % \Image{0.94\columnwidth}{images/week-8-m.jpg}
\end{center} 
\begin{itemize}
  \item 
\end{itemize}

\subsection{Penis With Penile Urethra}
\begin{center}
  % \Image{0.94\columnwidth}{images/week-8-n.jpg}
\end{center} 
\begin{itemize}
  \item 
\end{itemize}

\subsection{Questions}
\begin{itemize}\color{minor}
  \item Give an overview of the process of spermatogenesis and spermiogenesis.
  \basec{\begin{itemize}
    \item 
  \end{itemize}}

  \item What is the effect of castration on the accessory sex glands (prostate and seminal vesicles)?
  \basec{\begin{itemize}
    \item 
  \end{itemize}}

  \item Where is the primary source of testosterone?
  \basec{\begin{itemize}
    \item 
  \end{itemize}}

  \item Where is the principal site of storage of spermatozoa in the male reproductive system?
  \basec{\begin{itemize}
    \item 
  \end{itemize}}

  \item Follow the passage of spermatozoa from the seminiferous tubules of the testis up to ejaculation.
  \basec{\begin{itemize}
    \item 
  \end{itemize}}

  \item Which organs are the major sources of seminal fluid?
  \basec{\begin{itemize}
    \item 
  \end{itemize}}

  \item What are the components of the blood-testis barrier and what is its significance?
  \basec{\begin{itemize}
    \item 
  \end{itemize}}
\end{itemize}

