\chapter{22: Reactions of Aromatic Compounds}\label{22: Reactions of Aromatic Compounds}

\section{Electrophilic Aromatic Substitution}\label{Electrophilic Aromatic Substitution}
\begin{itemize}
  \item \ddd{Electrophilic Aromatic Substitution (EAS)}: a reaction where an atom (usually hydrogen) is replaced by an \rrr{electrophile} while the product retains aromaticity. 
  \item The electrophile typically used:
    \begin{itemize}
      \item \rrr{\ch{Br^{+}}} or \rrr{\ch{Cl^{+}}} (via Lewis acid catalyst) \smallskip
      \item Nitronium ion: \rrr{\ch{NO2^{+}}} \(\rightarrow \) \chemfig{O=\pab{N}=O}
      \item \rrr{\ch{SO3H^{+}}}
      \item \ang{2} or \ang{3} \rrr{\ch{R^{+}}} (not \ang{1})
      \item Acylium ion: \rrr{\ch{RCOX}}
      
      \medskip
      \schemestart{}
        \chemfig{R-[:30](=[:90]O)-[:-50,0.5,,,draw=none]\pab{}} 
        \arrow{<->}
        \chemfig{R-~\pab{O}}
      \schemestop{}
      \bigskip
    \end{itemize}
  
  \item \ddd{Arenium ion}: aka \(\sigma \)-complex (sigma complex); the reactive cyclohexadienyl cation that appears as a \rrr{reactive intermediate} for EAS\@.
  
  \bigskip
  \begin{center}
  \hspace{-30pt}
  \schemestart{}
    {\footnotesize\chemfig{*6(=-=-(-[1]H)(-[3]E)-(-[4,0.3,,,draw=none]\pab{})-)}}
    \arrow{<->}
    {\footnotesize\chemfig{*6(-(-[:-90,0.7,,,draw=none]\pab{})-=-(-[1]H)(-[3]E)-=)}}
    \arrow{<->}
    {\footnotesize\chemfig{*6(-=-(-[0,0.3,,,draw=none]\pab{})-(-[1]H)(-[3]E)-=)}}
  \schemestop{}
  \medskip

  \item Note: the hydrogen is then removed with a \bbb{base} in order to regain aromaticity.
  \end{center}
  
  \subsection{Halogenation of Benzene}\label{Halogenation of Benzene}
  \begin{itemize}
    \item Benzene can be chlorinated or brominated in the presence of a \rrr{Lewis acid catalyst}, typically ferric (iron III) chloride (\rrr{\ch{FeCl3}}) or ferric bromide (\rrr{\ch{FeBr3}}).
      \begin{itemize}
        \item The catalyst reacts with \ch{X_2} to form the strong electrophile; the catalyst regenerated by end of the reaction. 
        \item \ch{Fe.} may be used instead if used in the presence of \ch{Cl2} or \ch{Br2} since it will form the catalyst. 
        \item Aluminum chloride (\rrr{\ch{AlCl3}}) can be used as well, cost is only real factor.
      \end{itemize}

      \bigskip
      \begin{center}
      \hspace{-30pt}
      \schemestart{}
        {\footnotesize\chemfig{*6(-=-=-=)}}
        \arrow{->[\ch{Br2}][\ch{FeBr3}]}
        {\footnotesize\chemfig{*6(-=-=(-Br)-=)}}
      \schemestop{}
      \bigskip
      \end{center}

  \end{itemize}
  
  \subsection{Nitration of Benzene}\label{Nitration of Benzene}
  \begin{itemize}
    \item In benzene nitration, the electrophile is the \rrr{nitronium ion, \ch{NO2^{+}}}; it's formed by the protonation of nitric acid by the stronger sulfuric acid (\ch{H2SO4}), followed by loss of \ch{H2O}.
    \item Sulfuric acid is regenerated and recycled in the reaction.
    
    \bigskip
      \begin{center}
      \hspace{-30pt}
      \schemestart{}
        {\footnotesize\chemfig{*6(-=-=-=)}}
        \arrow{->[\ch{HNO3}][\ch{H2SO4}]}[,1.2]
        {\footnotesize\chemfig{*6(-=-=(-\ch{NO2})-=)}}
      \schemestop{}
      \end{center}
    \bigskip

    \item Once the nitro group is added, then a reduction step using Zinc (Zn) or Tin (Sn) and hydrochloric acid is done to make an aniline (benzene with \ch{NH2}).
  \end{itemize}
  
  \subsection{Sulfonation of Benzene}\label{Sulfonation of Benzene}
  \begin{itemize}
    \item In this reaction sulfuric acid protonates sulfur trioxide to form the \rrr{electrophilic sulfur-containing compound, \ch{SO3H^{+}}}. 
    \item Sulfuric acid is regenerated and recycled in the reaction.

  \bigskip
    \begin{center}
    \hspace{-30pt}
    \schemestart{}
      {\footnotesize\chemfig{*6(-=-=-=)}}
      \arrow{->[\ch{SO3}][\ch{H2SO4}]}[,1.2]
      {\footnotesize\chemfig{*6(-=-=(-\ch{SO3H})-=)}}
    \schemestop{}
    \end{center}
  \bigskip

  \end{itemize}
  
  \subsection{Friedel-Crafts Alkylation of Benzene}\label{Friedel-Crafts Alkylation of Benzene}
  \begin{itemize}
    \item \ddd{Friedel-Crafts alkylation (F.C. alkylation)}: alkylation of an aromatic ring with an alkyl halide using a strong Lewis acid catalyst, typically \rrr{aluminum chloride}, \rrr{ferric chloride}, but ferric bromide is also possible, similar to \hyperref[Halogenation of Benzene]{\ulink{halogenation of benzene}}.
    \item F.C alkylation reactions only work with methyl or ethyl groups; will not work correctly for longer groups, e.g., addition of a propyl alkyl becomes isopropyl substituent. 
    \item Specific structure of the electrophile can vary due to carbocation stability trends.
      \begin{itemize}
        \item \emph{Case 1 (MeX or EtX)}: a methyl or \ang{1}-carbocation cannot from and no rearrangements are possible; MeCl and EtCl complex with \ch{AlCl3} is the elctrophile.
        \item \emph{Case 2 (\ang{1} RX)}: the \ang{1} RCl complex with \ch{AlCl3} loses \bbb{\ch{AlCl4^{-}}} during the rearrangement, creating a \ang{2} or \ang{3} \rrr{\ch{R^{+}}}.
        \item\emph{Case 3 (\ang{2} or \ang{3} RX)}: the \ang{2} or \ang{3} RCl complex with \ch{AlCl3} loses \bbb{\ch{AlCl4^{-}}} directly creating the \ang{2} or \ang{3} \rrr{\ch{R^{+}}} (with rearrangement of \ang{2} \rrr{\ch{R^{+}}} occurring, if feasible)
      \end{itemize}
    \item The Lewis acid catalyst is regenerated and recycled during the reaction with the addition of \ch{HCl} byproduct much like halogenation.

    \bigskip
      \begin{center}
      \hspace{-30pt}
      \schemestart{}
        {\footnotesize\chemfig{*6(-=-=-=)}}
        \arrow{->[\ch{RCl}][\ch{AlCl3}]}[,1.2]
        {\footnotesize\chemfig{*6(-=-=(-\ch{R})-=)}}
      \schemestop{}
      \end{center}
    \bigskip

  \end{itemize}
  
  \subsection{Friedel-Crafts Acylation of Benzene}\label{Friedel-Crafts Acylation of Benzene}
  \begin{itemize}
      \item \ddd{Friedel-Crafts acylation}: when a new carbon-carbon bond is formed between a benzene carbon and a carbonyl carbon to create a ketone.
      \item Acylation is similar to alkylation, except:
        \begin{itemize}
          \item the electrophile is an \rrr{acylium ion} derived from an acid chloride;
          \item rearrangements will not occur because of the stability of the resonance-stabilized acylium ion.
        \end{itemize}

      \bigskip
        \begin{center}
        \hspace{-30pt}
        \schemestart{}
          {\footnotesize\chemfig{*6(-=-=-=)}}
          \arrow{->[\ch{RCOCl}][\ch{AlCl3}]}[,1.3]
          {\footnotesize\chemfig{*6(-=-=(-(=[:140]O)-[:30]\ch{R})-=)}}
        \schemestop{}
        \end{center}
      \bigskip

    \item Can be used to introduce more carbons to the alkyl group than alkylation alone can with later reduction techniques to remove the oxygen.
        \begin{itemize}
          \item \ddd{Clemmensen reduction}: zinc in presence of mercury (Zn(Hg)) with concentrated hydrochloric acid can be used in the reduction step.
          \item \ddd{Wolf-Kishner reduction}: hydrazine (\ch{N2H4}; \ch{NH2-NH2}) in the presence of a base and heat can also be used for reduction.
          \item Differences in pathways depends on substituents and their interaction with various reduction agents.
        \end{itemize}
    \item \emph{Intramolecular} F.C. acylation reactions can also occur to create \emph{fused rings}.
    \item Addition of \ch{KMNO4} with a strong base and heat acts as a strong oxidizing agent, resulting in any carbon attached directly to the ring becoming a carboxylic acid. 
  \end{itemize}

  \subsection{EAS Reactions of Disubstituted Benzenes}\label{EAS reactions of Disubstituted Benzenes}
  \begin{itemize}
      \item Agents and their products remain the same as monosubstituted benzenes, but the location of the new group can vary.
      \item \bbb{EDG push} electron density toward the benzene ring, making it more \bbb{electron rich} (\bbb{activated}).
        \begin{itemize}
          \item \bbb{\ch{-NH2}, \ch{-OH}, \ch{-OR}, \ch{-R}, \ch{-NHR}, \ch{-NR2}}
          \item \bbb{EDG} are \bbb{ortho/para} directing, with para preference increasing with increasing steric hindrance; small groups are approximately equal with slight ortho preference due to charge separation distance.
        \end{itemize}
      \item \rrr{EWG pull} electron density away from the benzene ring, making it \rrr{electron poor}, reducing electrophilic substitution tendency (\rrr{deactivated}).
        \begin{itemize}
          \item \rrr{\ch{-COOR}, \ch{-CONH2}, \ch{-CONR2}, \ch{-CN}, \ch{-NO2}, \ch{-COOH}, \ch{-CHO}, \ch{-COR}}
          \item \rrr{EWG} are \rrr{meta} directing, as the ortho/para positions remain more \rrr{electron poor}.  
        \end{itemize}
      \item When there are two (or more) substituents, then the \bbb{strongest activator} (the \bbb{strongest EDG} or \rrr{weakest EWG}) determines where the incoming electrophile reacts.
  \end{itemize}
  
  \subsection{Issues with EAS Reactions}\label{Issues with EAS Reactions}
  \begin{itemize}
    \item If the ring is \bbb{extremely activated} (typically with an \ch{-OH}), then the product results in a \emph{tribromophenol} product in presence of a Lewis acid catalyst.
    \begin{itemize}
      \item \ch{Br2} by itself can be used on a phenol to create a mono para/ortho product. 
    \end{itemize}
    \item If an \emph{aniline} (benzene + \ch{NH2}) undergoes \emph{halogenation} then the Lewis acid catalyst reacts with the \ch{NH2}, \emph{blocking the reaction}. 
      \begin{itemize}
        \item Not using a catalyst will allow the reaction, producing a mix of ortho/para products since \bbb{\ch{-NH2}} is a strong \bbb{EDG}.
      \end{itemize}
    \item \emph{F.C. alkylation} reactions do not work with a strong \rrr{EWG} that \rrr{deactivates} the ring.
  \end{itemize}
  
\end{itemize}