\documentclass[12pt,a4paper]{article}
\usepackage{inverba}

\newcommand{\userName}{Cullyn Newman} 
\newcommand{\class}{BI 358} 
\newcommand{\institution}{Portland State Univsrsity} 
\newcommand{\theTitle}{\color{G-Leaf}Evolution}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tableofcontents
\cleardoublepage
\fancyhead{}
\fancyhead[R]{\hyperlink{home}{\nouppercase\leftmark}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage
%\begingroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\section{Evolutionary Thinking}\phantomsection
\subsection{Introduction (Lecture)}
\begin{itemize}
    \item Essential questions of evolutionary biology: 
        \begin{itemize}
            \item Why do organisms look so different?
            \item Why develop elaborate sexual traits?
            \item Why do organisms senesce?
        \end{itemize}
    \item Evolution is mainly an historical science and thus must relay on other methods of reconstructing the past or making inferences about evolutionary forces. 
    \item \textbf{Proximate}: a question about a mechanistic cause; provides an immediate explanation about {\color{o-Sun}how} a mechanistic cause functions.
    \item \textbf{Ultimate}: {\color{o-Sun}why}, or the reason, a trait or organism is the way it is; an evolutionary explanation.
    \item Example of proximate vs ultimate in Galapagos finches: 
        \begin{itemize}
            \item Proxmate: developmental growth factor is increased/decreased in some birds.
            \item Ultimate: different habits are selected on breaks that maxmize food gathering ability.
        \end{itemize}
    \item Evolutionary biology's approach to answering questions:
        \begin{itemize}
            \item \textbf{Empirical data}: {\color{o-Sun}observation} studies, experiments; the \textit{comparative method}.
            \item \textbf{Theory}: {\color{o-Sun}predictions} that use models and mathematical reasoning which can be be {\color{o-Sun}tested} with empirical data.
        \end{itemize}
    \item Overview of the components of evolution by natural selection:
        \begin{itemize}
            \item Genetic variation exists, via mutations.
            \item Mutations are heritable.
            \item The is an advantage to survival and/or reproduction from the mutation.
            \item Individuals with the advantage in survival/reproduction are selected for.
        \end{itemize}
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 10 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\setcounter{section}{9}
\section{Studying Adaptation}\phantomsection
\subsection{Hypothesis Testing: Oxpeckers Reconsidered}
\begin{itemize}
    \item \textbf{Adaption}: a trait, or a suite of traits, that increases the fitness of its possessor.
    \item No hypothesis for the adaptive value of a trait should be accepted simply because of its plausibility.
    \item Oxpeckers and impalas traditionally were thought to have a mutually beneficial existence; oxpeckers ate ticks and impalas provided a safe environment.
    \item Experiments on cattle were done to test whether this observation was true:
        \begin{itemize}
            \item Results show red-billed oxpeckers have no effect tick loads of cattle.
            \item Red-billeld oxpeckers maintained open wounds, even enlarging existing wounds to feed on the cattle's blood.
            \item Red-billed oxpeckers removed hosts' earwax; whether this is good of bad is unclear.
            \item Even these results must remain in question, as cattle are not the native host for the birds.
        \end{itemize}
    \item Other important points to remember: 
        \begin{itemize}
            \item Differences among populations or species are not always adaptive.
            \item Not every trait is adaptive.
            \item Not every adaptation is perfect, often the adaption just happened to work well enough or by chance better than other adaptations.
        \end{itemize} 
\end{itemize}

\subsection{Experimental Design}
\begin{itemize}
    \item Defining and testing effective control groups is critical.
    \item Treatments of controls and experimental measures must be handled as close to exactly alike as possible.
    \item Randomization is a key technique for equalizing miscellaneous effects and a tool to avoid bias.
    \item Reproduction is essential in order to help remove potential outlier effects.
        \begin{itemize}
            \item Allows for greater understanding of precision, accuracy, and variation by providing more data for statistical tests.
        \end{itemize}
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 10 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 4 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\setcounter{section}{3}
\section{Evolutionary Trees}\phantomsection
\subsection{How to Read an Evolutionary Tree}
\begin{itemize}
    \item \textbf{Phylogeny}: aka evolutionary tree or phylogenetic tree, is a diaggram showing the history of divergence and evolutionary change. Essentially, it's the {\color{o-Sun}genealogical relationships} of organisms based on descent with modification.
        \begin{itemize}
            \item \textbf{Taxa}: the units you are analyzing, e.g. certain species or DNA sequences.
                \begin{itemize}
                    \item \textbf{Character}: a feature or trait present among the taxa of interest, e.g. teeth of mammals or nucleotides of DNA sequences.
                        \begin{itemize}
                            \item \textit{Character {\color{o-Sun}state(s)}}: an {\color{o-Sun}alternative condition} of a character, which are able to evolve one to another, e.g. pointed/flat teeth of mammals.
                        \end{itemize}
                    \item \textit{{\color{true}Ancestral} character}: a trait that was {\color{true}possessed by the common ancestor}.
                    \item \textit{{\color{false}Derived} character}: a trait the was {\color{false}not possessed by the common ancestor} and instead {\color{o-Sun}evolved} in at least one of the descendants.
                        \begin{itemize}
                            \item \textbf{Synapomorphy}: derived character state shared by {\color{o-Sun} two or more} taxa and used to define a clade of taxa.
                            \item \textbf{Autapomorphy}: derived character state in only {\color{o-Sun}one} taxon.
                        \end{itemize}
                    \item \textbf{Outgroup}: a taxon or taxa that are used to root the phylogeny or determine ancestral character states.
                    \item \textbf{Ingroup}: the set of taxa that are the focus of the phylogeny.
                \end{itemize}
            \item \textbf{Nodes}: points at which the tree splits; represents mutations, speciation events, or {\color{o-Sun}character changes}.
        \item {\color{false}\textbf{Anagensis}}: descent with modification, but {\color{false}no speciation}.
        \item {\color{true}\textbf{Cladogenesis}}: {\color{true}speciation}, origin of clades.
        \begin{itemize}
            \item \textbf{Clade}: also known as a {\color{o-Sun}monophyletic group}, an ancestor and {\color{o-Sun}all} of its descendants.
            \item \textbf{Paraphyletic group}: a group of organisms consisting of an ancestor and {\color{o-Sun}some} of its descendants.
        \end{itemize}
        \item \textbf{Sister}: a taxa or clade that are most closely related to each other; they {\color{o-Sun}share the most recent} common ancestor. 
    \end{itemize}
    \item \textbf{Homology}: similarity due to common descent; {\color{true}continuity} of a trait, character, or character state through time.
        \begin{itemize}
            \item \textit{Homologous trait}: found in a taxa that inherited the trait from a common ancestor.
        \end{itemize}
    \item \textbf{Homoplasy}: or analogous, similarity in the characters or traits in different taxa due to convergent evolution, parallelism, or reversal, but {\color{false}not due to common descent}.
        \begin{itemize}
            \item \textbf{Convergent evolution}: similar traits due to selective forces and {\color{false}not shared ancestry}.
                \begin{itemize}
                    \item \textbf{Parallelism}: convergent evolution in {\color{o-Sun}recently diverged} taxa.
                \end{itemize}
            \item \textbf{Reversal}: derived traits or character states that revert to the ancestral form.
        \end{itemize}
\end{itemize}

\subsection{Inferring Phylogenetic Trees}
\begin{itemize}    
    \item \textbf{Parsimony}: relationships that require the {\color{o-Sun}smallest number of character changes} are most likely to be correct.
        \begin{itemize}
            \item Based on shared and {\color{false}derived} traits(synapomorphies).
            \item Reconstruction using parsimony:
                \begin{itemize}
                    \item[1.] Code characters.
                    \item[2.] Make up a taxon$\times$character matrix.
                    \item[3.] Search for synapomorphies, and the shortest tree.
                    \item Outgroups can help polarize (ancestral vs derived) the characters. 
                \end{itemize}
            \item \textbf{Treelength}: a measure of evolutionary change using parsimony. 
                \begin{itemize}
                    \item Shortest tree length produces most parsimonious tree.
                    \item Length determined by number of synapomorphies.
                    \item Homoplasious characters increase tree length.
                \end{itemize}
        \end{itemize}
    \item \textbf{Distance Methods}: converts a sequence alignment to genetic distances between pairs of sequences.
        \begin{itemize}
            \item Branch length is proportional to genetic differences.
        \end{itemize}
    \item \textbf{Maximum likelihood}: a method of estimating the parameters of a probability distribution by {\color{o-Sun}maximizing a \textit{likelihood function}}. 
        \begin{itemize}
            \item One of the more dominant means of statistical inference.
            \item \textbf{Likelihood}: measure of goodness of fit of a statistical model to a sample of data for given values of the unknown parameters.
            \item {\color{o-Sun}\(P(D|H)\)}; probability(P), Data(D), Hypothesis(H)
            \item \textbf{Bayesian}: uses the likelihood function to create a quantity called the \textit{posterior probability} of trees using a model of evolution based on prior probabilities in order to produce the most likely tree.
            \item \textbf{Bootstraping}: creating a value that indicates how many times out of 100 (normally) that the same branch was observed when repeating the phylogenetic reconstruction on re-sampled (pseudoreplicated) set of dat.
        \end{itemize}
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 4 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 6 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\setcounter{section}{5}
\section{Mechanisms of Evolutionary Change}\phantomsection
\subsection{Hardy-Weinberg Equilibrium}
\begin{itemize}
    \item \textbf{Population}: a group of interbreeding individuals and their offsring.
    \item \textbf{Gene pool}: the set of all genes, or genetic information, in any population.
    \item \textbf{Genotypic frequency}: number of individuals with a given genotype divided by the total number of individuals in the population.
        \begin{itemize}
            \item The proportion (i.e., \(0<f<1\)) of genotypes in a population.
        \end{itemize}
    \item \textbf{Allele frequencies}: relative frequency of an allele at a particular locus in a population.
        \begin{itemize}
            \item \textbf{Locus}: a fixed position on a chromosome where a particular gene of genetic marker is.
            \item Monoploids: frequency of an allele is the result of the number of copies of the allele divided by sample size.
                \begin{itemize}
                    \item \(p=i/N\)
                    \item \(p\): frequency |  \(i\): copies of alleles |  \(N\): sample size
                \end{itemize}
            \item Diploids: frequency of alleles within three possbile genotypes at a locus with two alleles.
                \begin{itemize}
                    \item \({\color{o-Sun}p}=f({\color{o-Sun}AA})+\frac{1}{2}f(AB)\)~~frequency of {\color{o-Sun}A}-allele 
                    \item \({\color{o-Sun}q}=f({\color{o-Sun}BB})+\frac{1}{2}f(AB)\)~~frequency of {\color{o-Sun}B}-allele
                \end{itemize}
            \item Allele frequency can always be calculated from genotype frequency, whereas the reverse requires the \textit{Hardy-Weinberg principle} of random mating apply.
        \end{itemize}
    \item \textbf{Hardy-Weinberg principle}: allele and genotype {\color{o-Sun}frequencies} in a population will remain {\color{o-Sun}constant} in the {\color{o-Sun}absence of evolutionary influences}.
        \begin{itemize}
            \item Allele frequencies do not change from one generation to the next.
            \item Genotypic frequencies after one generation of random mating: \(p^2 + 2pq + q^2\)
            \item Evolutionary influences: genetic drift, mate choice, assortative mating, natural selection, sexual selection, mutation, gene flow, meiotic drive, genetic hitchhicking, population bottleneck, founder effect, and inbreeding.
                \begin{itemize}
                    \item \textit{Most of these influences will be discussed later}.
                \end{itemize}
        \end{itemize}
\end{itemize}

\subsection{Selection}
\begin{itemize}
    \item \textbf{Fitness}: success at which a organism produces fertile offspring.
    \item \textbf{Competition}: an interaction between organism in which the fitness of one is lowered by the presence of another.
    \item \textbf{Selection}: the act on a heritable phenotypic trait due to competition.
    \begin{itemize}
        \item Can be members of the same of different species.
        \item Not always directional and adaptive, instead selection pressure is applied and removes the less fit variants.
        \item Can be classified in different ways, such as effect on a trait, on genetic diversity, by life cycle, by unit of selection, or by the resource in competition.
        \item Most effective on large populations.
    \end{itemize}
    \subsubsection{By Effect on a Trait}
    \begin{itemize}
        \item \textbf{Stablizing selection}: the simplies case in which selection acts to hold a trait at a stable optimum.
        \item \textbf{Directional selection}: favours extreme values of a trait.
        \item \textbf{Disruptive selection}: acts during transition periods when current mode is sub-optimal, but alters trait in more than one direction.
            \begin{itemize}
                \item \textbf{Univariate}: when the trait is both quantitatively favoured in either direction and can lead to speciation.
            \end{itemize}
    \end{itemize}
    \subsubsection{By Effect on Genetic Diversity}
    \begin{itemize}
        \item \textbf{Purifying selection}: aka negative selection; acts to remove genetic variation from the population.
        \item \textbf{\textit{de novo} mutation}: introduces new variation and opposes negative selection.
        \item \textbf{Balancing selection}: acts to maintain genetic variation, even in absence of \textit{de novo} mutation by frequency-dependent selection.
            \begin{itemize}
                \item \textbf{Frequency-dependent selection}: fitness that depends of the phenotypic or genotypic {\color{o-Sun}composition} of a population.
                    \begin{itemize}
                        \item {\color{pos}Positive}: fitness {\color{pos}increases} as frequency of the trait {\color{pos}increases}.
                        \item {\color{neg}Negative}: fitness {\color{neg}decreases} as the frequency of the trait {\color{pos}increases}.
                    \end{itemize}
                \item \textbf{Overdominance}, \textit{aka heterozygote advantage}: when a combination of alleles confers a selective advantage over individuals with one allele.
                \item \textbf{Underdominance}, \textit{aka heterozygote disadvantage}: when the heterozygote has lower fitness than either homozygote. 
            \end{itemize}
    \end{itemize}
    \subsubsection{By Life Cycle Stage}
    \begin{itemize}
        \item \textbf{Viability selection}: \textit{aka survival selection}: increases probability of survival. 
            \begin{itemize}
                \item Can act to improve probability of survival before and after reproduction.
            \end{itemize}
        \item \textbf{Fecundity selection}: increases the rate of reproduction given survival.
            \begin{itemize}
                \item May be split into sub-components including sexual selection, gametic selection, gamete viability, compatability selection, and zygote formation.
            \end{itemize}
    \end{itemize}
\end{itemize}

\subsection{Mutation}
\begin{itemize}
    \item \textbf{Mutation}: alteration in the nucleotide sequence of the genome of an organism.
        \begin{itemize}
            \item May not produce discernible phenotypic changes.
            \item The ultimate source of genetic variation.
            \item Have several types of changes, from no effect, to small changes, or complete loss of function.
        \end{itemize}
    \subsubsection{Large-Scale Structural Mutations}
    \begin{itemize}
        \item \textbf{Gene duplications}, \textit{aka amplifications}: repetition of a chromosomal segment or attachment of extra piece of chromosome to another, leading to multiple copies of chromosomal regions.
        \item Deletions of large chromosmal regions.
        \item \textbf{Fusion genes}: mutations that join previously separated genes into one new distinct gene.
        \item \textbf{Chromosmal rearrangement}: large scale changes in structure of chromosomes, leading to speciation in isolated, inbred populations. Includes:
            \begin{itemize}
                \item \textbf{Chromosomal translocations}: interchange of genetic parts from nonhomologous chromosomes.
                \item \textbf{Chromosomal inversions}: reversing the orientation of a chromosomal segment.
                \item Non-homologous chromosomal crossover.
                \item \textbf{Interstitial deletions}: inverse of fusion genes; removes a segment of DNA joining distant genes.
            \end{itemize}
        \item \textbf{Loss of heterozygosity}: loss of one allele, by deletion or genetic recombination, in a organism that previously had two different alleles. 
    \end{itemize}
    \subsubsection{Small-Scale Mutations}
    \begin{itemize}
        \item \textbf{Point mutation}: a single nucleotide base change, that can result in a variety of effects.
        \item \textbf{Insertions}: add one or more extra nucleotides into the DNA.
            \begin{itemize}
                \item Usually caused by transposable elements, or errors during replication or repeating elements.
                \item Can cause\textit{reading frame shift}, possibly effecting how many codons are read, and thus altering the gene product.
            \end{itemize}
        \item \textbf{Deletions}: remove one or more nucleotides from the DNA.
            \begin{itemize}
                \item Also can cause a reading frame shift like insertions.
                \item Generally irreversible.
            \end{itemize}
        \item \textbf{Substitutions}: exchange of a single nucleotide for another.
            \begin{itemize}
                \item Often classified as transitions or transversions.
                \item Generally a purine (A-G) for a purine, or a pyrimidine (C-T) for a pyrimidine.
                \item Can be reversed by another point mutation.
            \end{itemize}
    \end{itemize}
    \subsubsection{Impact on Protein Sequence}
    \begin{itemize}
        \item Effect of mutation depends heavily on where it occurs, particularly in a coding or non-coding region.
        \item Regulator sequences, e.g. promoters, enhancers, silencers, can alter gene expression but are less likely to alter protein sequence.
        \item \textbf{Frameshift mutation}: caused by insertion or deletion of nucleotides that is not divisible by three, resulting in a different translation from the original.
        \item \textbf{Synonymous substitution}: a condon replacement with another that codes for same amino acid. 
            \begin{itemize}
                \item \textbf{Silent substitution}: no phenotypic difference after a synonymous substitution.
            \end{itemize}
        \item \textbf{Nonsynonymous substitution}: a codon replacement that codes for a different amino acid.
            \begin{itemize}
                \item \textbf{Missense mutation}: codon replacement that renders the resulting protein nonfunctional.
                \item \textbf{Nonsense mutation}: codon replacement that results in a premature stop codon that produces a truncated and often nunfunctional protein.
            \end{itemize}
    \end{itemize}
\end{itemize}
\subsection{Migration}
\begin{itemize}
    \item \textbf{Gene flow}: movement of alleles, or genetic variation, between populations.
        \begin{itemize}
            \item If the rate of gene flow is high enough, then two populations are considered to have equivalent allele frequencies and thus a single population.
            \item Constrains speciation by combining gene pools of the groups.
            \item May result in the addition of novel genetic variants in the gene pool.
        \end{itemize}
    \subsubsection{Factors of Gene Flow}
    \begin{itemize}
        \item Gene flow is expected to be lower in species that:
            \begin{itemize}
                \item have low mobility or dispersal.
                \item occur in fragmanted habits.
                \item have long distances between populations.
                \item have small population sizes.
            \end{itemize}
        \item \textbf{Allopatric speciation}: when gene flow is blocked by {\color{o-Sun}physical} barriers that inhibit gene flow.
        \item \textbf{Sympatric speciation}: result of gene flow that is blocked due to {\color{o-Sun}non-physical} barriers that inhibit gene flow.
            \begin{itemize}
                \item Often reproduction barriers are the main factors.
            \end{itemize}
    \end{itemize} 
\end{itemize}

\subsection{Genetic Drift}
\begin{itemize}
    \item \textbf{Genetic drift}: the change in the {\color{o-Sun} allele frequencies} in a population due to {\color{o-Sun}random sampling}.
        \begin{itemize}
            \item Not influenced by environmental factors.
        \end{itemize}
    \item May cause certain gene variants to become fixed or lost by chance.
    \item Generally drives populations towards genetic uniformity over time, {\color{o-Sun}decreasing heterozygosity}.
    \item Only mutation or gene flow can introduce new alleles, which acts against genetic drift.
    \item \textbf{Founder effect}: result of sampling error which has an increased likelyhood on populations with low numbers.
        \begin{itemize}
            \item By chance certain alleles can be dominant when they otherwise wouldn't be in a new founding population.
            \item Often acts to drastically increase rate of genetic drift.
        \end{itemize}
    \item \textbf{Genetic bottleneck}: a sharp reduction in the size of population due to environmental events. 
        \begin{itemize}
            \item Can essentially cause a founder effect, though it's not a new population.
        \end{itemize}
    \subsubsection{Coalescent Theory}
    \begin{itemize}
        \item \textbf{Coalescent theory}: how gene variants sampled from a population may have {\color{o-Sun}originated} from a common ancestor.
            \begin{itemize}
                \item Assumes no recombination, no natural selection, no gene flow in the simpiliest case.
            \end{itemize}
        \item Aims to look backward in time by merging allels into a single ancestral copy according to a random process in coalescence events.
        \item Many theoretical genealogies are made in order to compare to observed data in order to test assumptions about demographic history of a population.
            \begin{itemize}
                \item Used to make inference about population genetic parameters, such as migration, population size, and recombination.
            \end{itemize}
        \item \textbf{Coalescent time}: number of preceding generations where the coalescence took place, not calender time.
            \begin{itemize}
                \item Estimation of the time can be made multiplied by \(2N_e\) with the average time between generations.
                \item Time to coalescence for a pair of allels at a locus is {\color{o-Sun}dependent} on population size.
                \item Formula: \(P_c(t)=\left(1-\dfrac{1}{2N_e}\right)^{t-1}\left(\dfrac{1}{2N_e}\right)\)
            \end{itemize}
        \item Can also be used to model the amount of variation in DNA sequences expected from genetic drift and mutation.
    \end{itemize}
\end{itemize}

\subsection{Molecular Evolution}
\begin{itemize}
    \item \textbf{Molecular evolution}: the process of change in the sequence composision of cellular molecules across generations.
    \item \textbf{Polymorphism}: occurrence of two of more clearly different morphs, or alternative phenotypes, in the population of a species.
    \begin{itemize}
        \item \textbf{Substitution}: when allels become fixed or lost in a population and polymorphism is ended.
            \begin{itemize}
                \item Substitution rate (\(k\)): \(k = sN\mu\)
                \item \(s\) = probability of fixation.
                \item \(N\mu\)= mutation rate of population.
            \end{itemize}
    \end{itemize}
    \subsubsection{Recombination}
    \begin{itemize}
    \item \textbf{Recombination}: the process that results in genetic exchange between chromosomes or chromosomal regions.
        \begin{itemize}
            \item Can also cause mutations due to misalignment after recombination.
            \item \textbf{Gene repair}: a type of recombination that is the product of DNA repair that corrects damage using a homologous template. 
                \begin{itemize}
                    \item Often responsbile for homogenizing sequences of duplicate genes over long periods of time, which reduces nucleotide divergence.
                \end{itemize}
        \end{itemize}
    \item \textbf{Genetic hithicking}: change in allele frequency not because of natural selection, but due to proximity to a gene undering selective sweep.
        \begin{itemize}
            \item \textbf{Selective sweep}: a beneficial mutation that increases frequency and generally becomes fixed. 
        \end{itemize}
    \end{itemize}
    \subsubsection{Neutral Theory}
        \begin{itemize}
        \item \textbf{Neutral theory of molecular evolution}: most evolutionary changes occur at the molecular level.
        \item Most variation is due to random genetic drift of mutant alleles that are selectively neutral.
        \item Compatable with phenotypic evolution, as phenotypes are driven by molecular changes.
        \item Most mutations are neutral with respect to fitness.
        \item A minority of mutation are advantageous.
        \item Substitution rate predicted to be neutral, equal to per-individual mutation rate, {\color{o-Sun}independent} of population size. 
        \item \(K_A / K_s\) test used to determin direction selection based on evolutionary history.
            \begin{itemize}
                \item \(K_A\): number of nonsynonymous substitutions (replacement).
                \item \(K_S\): number of synonymous substitutions (silent).
                \item {\color{pos}\(K_A > K_s\)} signals for {\color{pos}positive selection}.
                \item {\color{neg}\(K_A < K_s\)} signals for {\color{neg}purifying selection}.
            \end{itemize}
        \end{itemize}
    \subsubsection{Molecular Clocks}
    \begin{itemize}
        \item \textbf{Molecular clocks}: the average rate at which species' genomes accumulates {\color{o-Sun}neutral mutations} over time.
        \begin{itemize}
            \item A linear rate is often easy to establish.
            \item Used to measure evolutionary divergence.
        \end{itemize}
    \end{itemize}
\end{itemize}

\subsection{Nonrandom Mating}
\begin{itemize}
    \item \textbf{Inbreeding}: production of offspring from closely genetically related individuals.
        \begin{itemize}
            \item Results in homozygosity, which can increase chances chances of offspring being affected by deleterious or recessive traits.
            \item \textbf{Inbreeding depression}: the reduced fitness in a given population due to inbreeding.
                \begin{itemize}
                    \item Usually caused by population bottlenecks or the founder effect.
                \end{itemize}
            \item Can also result in purging of deleterious allels through purifying selection.
            \item Can allow for the expression of advantageous phenotypes, which if outweighs the disadvantages, then could potentially lead to speciation.
    \item \textbf{Coefficient of inbreeding}: the probability that two alleles at any locus in an individual are identical by descent.
    \item Nonrandom mating does not alter allele frequencies and not a mechanism of evolution.
        \begin{itemize}
            \item Can alter the frequencies of genotypes, changing the distribution of phenotypes in a population, which can alter patters of natural selection.
        \end{itemize}
    \end{itemize}
    \item \textbf{Assortative mating}: mating based on phenotypic factors.
        \begin{itemize}
            \item Can play a role in sympatric speciation.
            \item A form of sexual selection.
            \item Can be either positive of negative, selecting for similar or different phenotypes respectively.
        \end{itemize}
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 7 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\endgroup

\clearpage
%\begingroup


%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 8 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\setcounter{section}{7}
\section{Evolution at Multiple Loci}\phantomsection
\subsection{Linkage Equilibrium and Diesquilibrium}
\begin{itemize}
    \item \textbf{Linkage equilibrium}: when the genotype of a chromosome at one locus is {\color{o-Sun}independent} of its genotype at another locus.
    \item \textbf{Linkage disequilibrium}: the non-random association of alleles at different loci in a given population. 
        \begin{itemize}
            \item Occurs when frequency of the association between loci's different alleles is higher or lower than expected.
        \end{itemize}
    \item \textbf{Haplotype}: a group of alleles in an organism that are inherited together from a single parent.
        \begin{itemize}
            \item Used to mean the collection of specific alleles that represent a phenotype and likely to be conserved.
            \item Also can be used to mean a set of linked single-nucleotide polymorphism alleles that are associated statistically.
        \end{itemize}
    \item Factors that influences disequilibrium: selection, rate or genetic recombination, mutation rate, genetic drift, system of mating, population structure.
        \begin{itemize}
            \item Undertanding linkage desequilibrium in a genome can be a powerful signal of the population genetic processes that structure it.
            \item Selection, genetic drift, assortative mating, and population admixture act to {\color{pos}create disequilibrium}. 
            \item Recombination and outbreeding act to {\color{neg}reduce disequilibrium}.
        \end{itemize}
    \item Level of linkage disequilibrium can between \(A\) and \(B\) can be quantified by the {\color{o-Sun}coefficient of linkage disequilibrium, \(D_{AB}\)}.
        \begin{itemize}
            \item Formula: \(D_{AB}=P_{AB} - P_{A}P_{B}\)
            \item \(P_{AB}\): the frequency with which both occur together on same gamete, or the frequency of the \(AB\) haplotype.
            \item \(P_AP_B\): product of the probabilities give the probability they occur together.
            \item When there is a difference, the magnitude of the coefficient rises, indicating linkage disequilibrium.
            \item Strong recent selection can be indicated by linkage disequilibrium of allels located next to neutral allels.
        \end{itemize}
    \subsubsection{Recombination's Effect on  Linkage Diesquilibrium}
    \begin{itemize}
        \item Linkage diesquilibrium (\(D\)) will converge to zero depending of the magnitude of the recombination rate (\(c\)) between two loci the absence of natural selection, inbreeding, and genetic drift.
        \item The smaller the distance between the two loci, the smaller the rate of convergence of D to zero.
        \item Genetic recombination tends to randomize genotypes, thus it tends to reduce frequency of overrespresented chromosome haplotypes and increases underrepresented haplotypes. 
        \item In short, recombination reduces linkage disequilibrium.
    \end{itemize}
\end{itemize}

\subsection{Adaptive Significance of Sex}
\begin{itemize}
    \item 
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 8 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 9 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\section{Quantitative Genetics}\phantomsection
\subsection{The Nature of Quantitative Traits}
\begin{itemize}
    \item \textbf{Quantitative genetics}: deals with phenotypes that vary continuosly due to multilocus traits and environmental factors.
        \begin{itemize}
            \item Allows for the prediction of how a population will respond to selection, even when we do not know the genetic basis of the trait.
        \end{itemize}
    \item \textbf{Qualitative traits}: descrete traits that can be assigned individual categories by observation of simple genetic test.
    \item \textbf{Quantitative traits}: continuous traits determined by the combined influence of the genotype at multiple loci and the environment; the focus of quantitative genetics.
        \begin{itemize}
            \item Study of continuous distribution requires many other statistical methods such as the effect size, mean, and variance, to link phenotypes to genotypes.
        \end{itemize}
\end{itemize}

\subsection{Heritable Variation}
\begin{itemize}
    \item \textbf{Heritability}: the degree of variation in a phenotypic trait in a population that is due to genetic variation betwen individuals in that population.
    \item \textbf{Phenotypic variance (V\(_{P}\))}: the genetic variance (V\(_{G}\)) combined with the environmental variance (V\(_{E}\)).
        \begin{itemize}
            \item {\color{o-Sun}\(V_P = V_G + V_E + V_{GE}\)}
            \item \(V_{GE}\) represents variance associated with {\color{o-Sun}intereaction} of genetic and environmental factors. 
        \end{itemize}
    \subsubsection{Genetic Variance}
    \begin{itemize}
        \item \textbf{Additive genetic variance (V\(_{A}\))}: how much the phenotypic trait is influenced by traits that show an additive effect on the quantitative traits.
            \begin{itemize}
                \item Measures the magnitude to which individual phenotypic differences can be predicted due to additive effects of allelic substitutions.
            \end{itemize}
        \item \textbf{Dominance genetic variance (V\(_{D}\))}: associated with the dominant gene actions which cover the influence of the recessive alleles at the particular locus.
        \item \textbf{Epistasis (V\(_{I}\))}: occurs due to statistical interaaction among loci, i.e., gene-by-gene modification.
        \item  {\color{o-Sun}\(V_G = V_A + V_D + V_I\)}
    \end{itemize}
    \item \textbf{Broad-sense heritability (H\(^{2}\))}: all genetic contributions to a populations phenotypic variance, including additive, dominant, and epistatic, and maternal/paternal effects. 
        \begin{itemize}
            \item \(H^2 = \dfrac{V_G}{V_P}\)
        \end{itemize}
    \item \textbf{Narrow-sense heritability (h\(^{2}\))}: proportion of the total phenotypic varian that is due to the {\color{o-Sun}additive effects} of genes.
        \begin{itemize}
            \item \(h^2 = \dfrac{V_A}{V_P}\)
            \item Allows prediction of how a population will respond to selection.
            \item The greater the additive genetic variation for the trait, the greater its response to selection can be.
        \end{itemize}
\end{itemize}

\subsection{Measuring Differences in Survival and Reproductive Success}
\begin{itemize}
    \item Heritability is often estimated by measuring correlations between parents and offspring.
\end{itemize}

\subsection{Predicting Evolutionary Responses}
\begin{itemize}
    \item 
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 9 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\endgroup

\clearpage
%\begingroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 11 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\setcounter{section}{10}
\clearpage
\section{}\phantomsection
\subsection{}
\begin{itemize}
    \item 
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 11 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 12 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\section{}\phantomsection
\subsection{}
\begin{itemize}
    \item 
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 12 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 13 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\section{}\phantomsection
\subsection{}
\begin{itemize}
    \item 
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 13 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 14 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\section{}\phantomsection
\subsection{}
\begin{itemize}
    \item 
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 14 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 16 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\setcounter{section}{15}
\section{}\phantomsection
\subsection{}
\begin{itemize}
    \item 
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 16 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\endgroup

\clearpage
%\begingroup

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 17 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\section{}\phantomsection
\subsection{}
\begin{itemize}
    \item 
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 17 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 18 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\section{}\phantomsection
\subsection{}
\begin{itemize}
    \item 
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 18 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\endgroup

\clearpage
%\begingroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 20 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\setcounter{section}{19}
\section{}\phantomsection
\subsection{}
\begin{itemize}
    \item 
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 20 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 15 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begingroup
\clearpage
\setcounter{section}{14}
\section{}\phantomsection
\subsection{}
\begin{itemize}
    \item 
\end{itemize}
%\endgroup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 15 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\endgroup
\end{document}