\chapter{Limits and Continuity}

\section{Limits}

\src{\link{https://en.wikipedia.org/wiki/Limit_(mathematics)}{Limit} | \link{https://book4you.org/s/0134438981}{Thomas' Calculus (2.2--2.4)}}

\smallskip

\begin{itemize}
  \item \ddd{Limit \(\lim_{x \to c}\)}: the value of a function \prn{or sequence} approaches as the input \prn{or index} approaches some value \prn{informal definition}. 
  \begin{itemize}
    \item Limits are used to define \hyperref[s: Continuity]{\dlink{continuity}}, \hyperref[c: Derivatives]{\dlink{derivatives}}, and \hyperref[c: Integrals]{\dlink{integrals}}.
  \end{itemize}

  \subsection{Limits of a Functions and Sequences}
  \src{\link{https://en.wikipedia.org/wiki/Limit_of_a_function}{Limit of a function} | \link{https://en.wikipedia.org/wiki/Limit_of_a_sequence}{Limit of a sequence} | \linky{https://youtu.be/kfF40MiS7zA?t=281}{Essence of Calculus, E7}}
  \begin{itemize}
    \item \ddd{Limit of a function}: a fundament concept in calculus and analysis concerning the behavior \(L\) of a function near a particular input \(\yyy{c}\), i.e., 
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \lim_{\x \to\yyy{c}}f(\x) = L
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{itemize}
      \item Reads as ``\(\f\) of \(\x\) tends to L as \(\x\) tends to \(\yyy{c}\)''
    \end{itemize}
    \item \ddd{\(\epsilon,\delta\) Limit of function}: a formalized definition, wherein \(f(x)\) is defined on an open interval \(\bBb{I}\), except possibly at \yyy{c} itself, leading to above definition, \ssec{if and only if}:
      \begin{itemize}
        \item For every real measure of \bbb{closeness} \(\bbb{\epsilon > 0}\), there exists a real \rrr{corresponding} \(\rrr{\delta >0}\), such that for all existing further approaches there exist a smaller \(\bbb{\epsilon}\), i.e.,
        \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        f : \R \to \R,~ \yyy{c}, L \in \R \Rightarrow \lim_{\x \to\yyy{c}}f(\x) = L 
        \]%%%%%%%%%[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \vspace{-24pt}
        \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \ssec{\Updownarrow}
        \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \forall\bbb{\depsilon >0} \left(
          \exists\rrr{\delta >0} : 
            \forall \x \in \bBb{I} \left(
              0 < | \x - \yyy{c} | < \rrr{\delta} 
              \Rightarrow 
              | f(\x) - L | < \bbb{\depsilon}
            \right)
        \right)
        \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        
      \end{itemize}
    \item Functions do not have a limit when the function:
      \begin{itemize}
        \item has a unit step, i.e., it ``jumps'' at a point;
        \item is not bounded, i.e., it tends towards infinity;
        \item or does not stay close to any single number, i.e., it oscillates too much. 
      \end{itemize}
    \item \ddd{Limit of a sequence}: the value that the terms of a sequence (\(x_\ppp{n}\)) ``tends to'' \prn{and not to any other} as \(\ppp{n}\) approaches infinity \prn{or some point}, i.e.,
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \lim_{\ppp{n} \to \yyy{\infty}} x_\ppp{n} = x
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      \item \ddd{\(\epsilon \) Limit of sequence}: for every measure of closeness \(\bbb{\epsilon} \), the sequence's term eventually converge to the limit, i.e.,
      \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      \forall\bbb{\depsilon > 0} \left(
        \exists \ppp{N} \in \N \left(
          \forall\ppp{n} \in \N \left(
            \ppp{n} \geq \ppp{N} \Rightarrow |x_\ppp{n}-x| < \bbb{\depsilon}
          \right)
        \right)
      \right)
      \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \begin{itemize}  
      \item \ddd{Convergent}: when a limit of a sequence \ttt{exists}.
      \item \ddd{Divergent}: a sequence that \fff{does not} converge. 
    \end{itemize}
  \end{itemize}

  \subsection{Properties of Limits}
  \src{\link{https://en.wikipedia.org/wiki/List_of_limits}{List of limits} | \link{https://en.wikipedia.org/wiki/Squeeze_theorem}{Squeeze theorem}}
  \begin{itemize}
    \item \ddd{Operations on a single known limit}: if \(\lim_{\x \to \yyy{c}}f(\x)=L\) then:
    \vspace{-6pt}
      \begin{itemize}
        \item \(\lim_{\x \to \yyy{c}}[f(\x)\pm \alpha] = L \pm \alpha\)  
        \item \(\lim_{\x \to \yyy{c}}\alpha f(\x)= \alpha L\)
        \item \(\lim_{\x \to \yyy{c}}f(\x)^{-1}=L^{-1}, L\neq 0\)
        \item \(\lim_{\x \to \yyy{c}}f(\x)^n = L^n, n \in \N\)
        \item \(\lim_{\x \to \yyy{c}}f(\x)^{n ^{-1}} = L^{n ^{-1}}, n \in \N,~\text{if}~n \in \N_e \then L > 0\) 
      \end{itemize}
    \medskip
    \item \ddd{Operations on two known limits}: if \(\lim_{\x \to \yyy{c}}f(\x)= L_1\) and \(\lim_{\x \to \yyy{c}}g(\x)=L_2\) 
    \vspace{-6pt}
      \begin{itemize}
        \item \(\lim_{\x \to \yyy{c}}[f(\x)\pm g(\x)]=L_1\pm L_2\)
        \item \(\lim_{\x \to \yyy{c}}[f(\x)g(\x)] = L_1 L_2\)
        \item \(\lim_{\x \to \yyy{c}}f(\x)g(\x)^{-1} = L_1L_2^{-1}\)
      \end{itemize}
    \medskip
    \item \ddd{Squeeze theorem}: used to confirm the limit of a function via comparison with two other functions whose limits are easily known or computed.
      \begin{itemize}
        \item Let \(\bBb{I}\) be an interval having the point \(\yyy{a}\) as a limit point. 
        \item Let \(g,f\), and \(h\), be functions defined on \(\bBb{I}\), except possibly at \(\yyy{a}\) itself. 
        \item Suppose that \(\forall \x \in \bBb{I}~\land~\x \neq \yyy{a} \then \bbb{g(\x)}\leq f(\x) \leq \rrr{h(\x)}\)
        \item And suppose that \(\lim_{\x \to \yyy{a}}\bbb{g(\x)}=\lim_{\x \to \yyy{a}}\rrr{h(\x)} = L\)
        \item Then, \(\lim_{\x \to \yyy{a}}f(\x) = L\)
        \item Essentially, the hard to compute limit of the ``middle function'' is found by finding two other easy functions that ``squeeze'' the middle function at that point.
      \end{itemize}
  \end{itemize}

  \subsection{One-Sided Limit}
  \src{\link{https://en.wikipedia.org/wiki/One-sided_limit}{One-Sided Limit}}
  \begin{itemize}
    \item \ddd{One-sided limit}: one of two limits of \(f(\x)\) as \(\x\) approaches a specified point from either the \bbb{left} or from the \rrr{right}.
    \begin{multicols}{2}
      \begin{itemize}
        \item From the \bbb{left}: \(\lim_{\x \to \bbb{c^-}} = L\)
        \item From the \rrr{right}:  \(\lim_{\x \to \rrr{c^+}} = L\)
      \end{itemize}
    \end{multicols}
    \item If the left and right limits exist and are equal, then 
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \lim_{\x \to \yyy{c}}f(\x) = L~\ssec{\Leftrightarrow}~ 
    \lim_{\x \to \bbb{c^-}}f(\x) = L ~\land
    \lim_{\x \to \rrr{c^+}}f(\x) = L
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \item Limits can still exist, even if the function is defined at a different point, as long as both one-sided limits approach the same value near the given input.
  \end{itemize}
  
\end{itemize}

\section{Continuity}
\src{
  \link{https://en.wikipedia.org/wiki/Continuous_function}{Continuous function} | 
  \link{https://en.wikipedia.org/wiki/Classification_of_discontinuities}{Classification of discontinuities} | 
  \link{https://book4you.org/s/0134438981}{Thomas' Calculus (2.5)}
}
\begin{itemize}
  \item Continuity of functions is one of the core concepts of topology, however, there are definitions in terms of limits that prove useful; the following is only a primer.

  \subsection{Continuous Functions}
  \begin{itemize}
    \item \ddd{Continuous function}: a function that does not have any abrupt changes in value.
      \begin{itemize}
        \item I.e., a function is continuous if and only if arbitrarily small changes in its output can be assured by restricting to sufficiently small changes in its input.
      \end{itemize}
    \item \ddd{Discontinuous}: when a function is not continuous at a point in its domain, leading to a discontinuity; there are three classifications:
      \begin{itemize}
        \item \ddd{Removable}: when both \hyperref[ss: One-Sided Limit]{\ulink{one-sided limits}} exist, are finite, and are equal, but the actual value of \(f(\x)\) is not equal to the limit and equal to some other value.
          \begin{itemize}
            \item The discontinuity can be removed to regain continuity.
            \item Sometimes the term \textit{removable discontinuity} is mistaken for \textit{removable singularity}, or a ``whole'' in the function \prn{the point is not defined elsewhere}.
          \end{itemize}
        \item \ddd{Jump}: when a single limit does not exist because the one-sided limits exist and are finite, but not equal.
          \begin{itemize}
            \item Points can be defined at the discontinuity, but the function can not be made continuous.
          \end{itemize}
        \item \ddd{Essential}: when at least one of two one-sided limits doesn't exist; can be the result of oscillating or unbounded functions.
      \end{itemize}
  \end{itemize}
  
  \subsection{Intermediate Value Theorem}
  \src{\link{https://en.wikipedia.org/wiki/Intermediate_value_theorem}{Intermediate Value Theorem}}
  \begin{itemize}
    \item \ddd{Intermediate value theorem}: if \(f\) is a continuous function whose domain contains the interval \([a,b]\), then it \emph{takes on any given value between} \(f(a)\) and \(f(b)\) at some point within the intervals.
    \item Relevant deductions, i.e., important corollaries:
      \begin{itemize}
        \item \ddd{Bolzano's theorem}: if a continuous function has values of opposite sign inside an interval, then it \emph{has a root} in that interval.
        \item The image of a continuous function over an interval is itself an interval. 
      \end{itemize}
    \item Thus, the image set \(f(\bBb{I})\) \prn{which has no gaps} is also an interval, and it contains: 
    \[%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \left[
      \min{f(a),f(b)},~
      \max{f(a),f(b)}
    \right]
    \]%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \end{itemize}
\end{itemize}

\section{Limits Involving Infinity}
\src{\link{https://en.wikipedia.org/wiki/Limit_of_a_function\#Limits_involving_infinity}{Limits involving infinity} | \link{https://book4you.org/s/0134438981}{Thomas' Calculus (2.6)} }
\begin{itemize}
  \item Let \(S \subseteq, \x \in S\) and \(f : S \mapsto \R\), then limits of these functions can approach arbitrarily large \prn{\(\pm\)} values, providing a connection to asymptotes, and thus, analysis.
 

  \subsection{Limits at Infinity}
  \begin{itemize}
    \item 
  \end{itemize}

  \subsection{Infinite Limits}
  \begin{itemize}
    \item 
  \end{itemize}
  
  \subsection{Asymptotes of functions}
  \src{\link{https://en.wikipedia.org/wiki/Asymptote}{Asymptote}}
  \begin{itemize}
    \item  \ddd{Asymptote}:  
  \end{itemize}

\end{itemize}






